---
title: "R for Data Science"
subtitle: "11. Data import"
author: Franciele Pereira Rossini 
date: "`r format(Sys.time(), '%d de %B de %Y')`"
output: 
   pdf_document:
     toc: true
     toc_depth: 2
---

```{r setup, include=FALSE}
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
```

# 11.1 Introduction

- Working with data provided by R packages is a great way to learn the tools of data science, but at some point you want to stop learning and start working with your own data. In this chapter, you'll learn how to read plain-text rectangular files into R. Here, we'll only scratch the surface of data import, but many of the principles will translate to other forms of data. We'll finish with a few pointers to packages that are useful for other types of data.
 
 
# 11.1.1 Prerequisites

- In this chapter, you’ll learn how to load flat files in R with the \textcolor{red}{readr package}, which is part of the core tidyverse.



# 11.2 Getting started

Most of readr's functions are concerned with turning flat files into data frames:

  - **read_csv()** reads comma delimited files;
    
  - **read_csv2()** reads semicolon separated files (common in countries where **,** is used as the decimal place);
    
  - **read_tsv()** reads tab delimited files; and 
    
  - **read_delim()** reads in files with any delimiter.

  - **read_fwf()** reads fixed width files.
     
    You can specify fields either by their widths with **fwf_widths()** or their position with **fwf_positions()**.
    
  - **read_table()** reads a common variation of fixed width files where columns are separated by white space.

  - **read_log()** reads Apache style log files.
  
These functions all have similar syntax: once you’ve mastered one, you can use the others with ease. For the rest of this chapter we’ll focus on read_csv(). Not only are csv files one of the most common forms of data storage, but once you understand read_csv(), you can easily apply your knowledge to all the other functions in readr.

The first argument to read_csv() is the most important: it’s the path to the file to read.


```{r, ex. 1}

mtcars<- read_csv (readr_example ("mtcars.csv" ))
mtcars
```


When you run read_csv() it prints out a column specification that gives the name and type of each column. That’s an important part of readr.


You can also supply an inline csv file. This is useful for experimenting with readr and for creating reproducible examples to share with others:

```{r, ex. 2}

read_csv("a,b,c
1,2,3
4,5,6")

```

In both cases read_csv() uses the first line of the data for the column names, which is a very common convention. There are two cases where you might want to tweak this behaviour:

  **1.** Sometimes there are a few lines of metadata at the top of the file. You can use **skip = n** to skip the first n lines; or use **comment = "#"** to drop all lines that start with (e.g.) **#**.


```{r, ex. 3}
read_csv("The first line of metadata
  The second line of metadata
  x,y,z
  1,2,3", skip = 2)
```

```{r, ex. 4}
read_csv("# A comment I want to skip
  x,y,z
  1,2,3", comment = "#")
```

  **2.** The data might not have column names. You can use **col_names = FALSE** to tell read_csv() not to treat the first row as headings, and instead label them sequentially from **X1 to Xn**:

```{r, ex. 5}
read_csv("1,2,3\n4,5,6", col_names = FALSE)

#\n is a convenient shortcut for adding a new line.
```

Alternatively you can pass col_names a character vector which will be used as the column names:

```{r, ex. 6}
read_csv("1,2,3\n4,5,6", col_names = c("x", "y", "z"))
```


Another option that commonly needs tweaking is na: this specifies the value (or values) that are used to represent missing values in your file:

```{r, ex. 7}
read_csv("a,b,c\n1,2,.", na = ".")
```


## 11.2.1 Compared to base R


If you’ve used R before, you might wonder why we’re not using read.csv(). There are a few good reasons to favour readr functions over the base equivalents:

- They are typically much faster (~10x) than their base equivalents. Long running jobs have a progress bar, so you can see what’s happening. If you’re looking for raw speed, try **data.table::fread()**. It doesn’t fit quite so well into the tidyverse, but it can be quite a bit faster.

- They produce tibbles, they don’t convert character vectors to factors, use row names, or munge the column names. These are common sources of frustration with the base R functions.

- They are more reproducible. Base R functions inherit some behaviour from your operating system and environment variables, so import code that works on your computer might not work on someone else’s



# 11.3 Parsing a vector

- **parse_*()** These functions take a character vector and return a more specialised vector like a logical, integer, or date: 

```{r, ex. 8}
str(parse_logical(c("TRUE", "FALSE", "NA")))

str(parse_integer(c("1", "2", "3")))

str(parse_date(c("2010-01-01", "1979-10-14")))

```
Like all functions in the tidyverse, the **parse_*()** functions are uniform: the first argument is a character vector to parse, and the na argument specifies which strings should be treated as missing:

```{r, ex. 9}
parse_integer(c("1", "231", ".", "456"), na = ".")

```

If parsing fails, you’ll get a warning:


```{r, ex. 10}
x <- parse_integer(c("123", "345", "abc", "123.45"))

x
```

If there are many parsing failures, you’ll need to use **problems()** to get the complete set. This returns a tibble, which you can then manipulate with dplyr.

```{r, ex. 11}
problems(x)
```
Using parsers is mostly a matter of understanding what’s available and how they deal with different types of input. There are eight particularly important parsers:

1. **parse_logical()** and **parse_integer()** parse logicals and integers respectively. There’s basically nothing that can go wrong with these parsers so I won’t describe them here further.

2. **parse_double()** is a strict numeric parser, and parse_number() is a flexible numeric parser. These are more complicated than you might expect because different parts of the world write numbers in different ways.

3. **parse_character()** seems so simple that it shouldn’t be necessary. But one complication makes it quite important: character encodings.

4. **parse_factor()** create factors, the data structure that R uses to represent categorical variables with fixed and known values.

5. **parse_datetime()**, **parse_date()**, and **parse_time()** allow you to parse various date & time specifications. These are the most complicated because there are so many different ways of writing dates.

##  11.3.1 1.3.1 Numbers

It seems like it should be straightforward to parse a number, but three problems make it tricky:

People write numbers differently in different parts of the world. For example, some countries use . in between the integer and fractional parts of a real number, while others use ,.

Numbers are often surrounded by other characters that provide some context, like “$1000” or “10%”.

Numbers often contain “grouping” characters to make them easier to read, like “1,000,000”, and these grouping characters vary around the world.

To address the first problem, readr has the notion of a “locale”, an object that specifies parsing options that differ from place to place. When parsing numbers, the most important option is the character you use for the decimal mark. You can override the default value of . by creating a new locale and setting the decimal_mark argument:

```{r, ex. 12}
parse_double("1.23")

parse_double("1,23", locale = locale(decimal_mark = ","))
```


readr’s default locale is US-centric, because generally R is US-centric (i.e. the documentation of base R is written in American English). An alternative approach would be to try and guess the defaults from your operating system. This is hard to do well, and, more importantly, makes your code fragile: even if it works on your computer, it might fail when you email it to a colleague in another country.


**parse_number()** addresses the second problem: it ignores non-numeric characters before and after the number. This is particularly useful for currencies and percentages, but also works to extract numbers embedded in text.


```{r, ex. 13}
parse_number("$100")

parse_number("20%")

parse_number("It cost $123.45")

```

The final problem is addressed by the combination of parse_number() and the locale as parse_number() will ignore the “grouping mark”:

```{r, ex. 14}
# Used in America
parse_number("$123,456,789")


# Used in many parts of Europe
parse_number("123.456.789", locale = locale(grouping_mark = "."))


# Used in Switzerland
parse_number("123'456'789", locale = locale(grouping_mark = "'"))

```

## 11.3.2 Strings

 In R, we can get at the underlying representation of a string using **charToRaw()**:


```{r, ex.15}
charToRaw("mtcars")
```



Each hexadecimal number represents a byte of information: 48 is H, 61 is a, and so on. The mapping from hexadecimal number to character is called the encoding, and in this case the encoding is called ASCII. ASCII does a great job of representing English characters, because it’s the American Standard Code for Information Interchange.

Things get more complicated for languages other than English. In the early days of computing there were many competing standards for encoding non-English characters, and to correctly interpret a string you needed to know both the values and the encoding. For example, two common encodings are **Latin1 (aka ISO-8859-1, used for Western European languages)** and **Latin2 (aka ISO-8859-2, used for Eastern European languages)**. In Latin1, the byte b1 is “±”, but in Latin2, it’s “ą”! Fortunately, today there is one standard that is supported almost everywhere: UTF-8. UTF-8 can encode just about every character used by humans today, as well as many extra symbols (like emoji!).


Readr uses UTF-8 everywhere: it assumes your data is UTF-8 encoded when you read it, and always uses it when writing. This is a good default, but will fail for data produced by older systems that don’t understand **UTF-8**. If this happens to you, your strings will look weird when you print them. Sometimes just one or two characters might be messed up; other times you’ll get complete gibberish. For example:


```{r, ex. 16}
x1 <- "El Ni\xf1o was particularly bad this year"
x1
x2 <- "\x82\xb1\x82\xf1\x82\xc9\x82\xbf\x82\xcd"
x2

```

To fix the problem you need to specify the encoding in **parse_character()**:

```{r, ex. 17 }

parse_character(x1, locale = locale(encoding = "Latin1"))

parse_character(x2, locale = locale(encoding = "Shift-JIS"))

```

How do you find the correct encoding? If you’re lucky, it’ll be included somewhere in the data documentation. Unfortunately, that’s rarely the case, so readr provides **guess_encoding()** to help you figure it out. It’s not foolproof, and it works better when you have lots of text (unlike here), but it’s a reasonable place to start. Expect to try a few different encodings before you find the right one.

```{r, ex. 18}

guess_encoding(charToRaw(x1))


guess_encoding(charToRaw(x2))
```

The first argument to **guess_encoding()** can either be a path to a file, or, as in this case, a raw vector (useful if the strings are already in R).

## 11.3.3 Factors

R uses factors to represent categorical variables that have a known set of possible values. Give **parse_factor()** a vector of known levels to generate a warning whenever an unexpected value is present:

```{r, ex. 19}

fruit <- c("apple", "banana")
parse_factor(c("apple", "banana", "bananana"), levels = fruit)

```

But if you have many problematic entries, it’s often easier to leave as character vectors and then use the tools you’ll learn about in strings and factors to clean them up.


## 11.3.4 Dates, date-times, and times

You pick between three parsers depending on whether you want a date (the number of days since 1970-01-01), a date-time (the number of seconds since midnight 1970-01-01), or a time (the number of seconds since midnight). When called without any additional arguments:

**parse_datetime()** expects an ISO-8601 date-time. ISO-8601 is an international standard in which the components of a date are organised from biggest to smallest: year, month, day, hour, minute, second.


```{r, ex. 20}
parse_datetime("2010-10-01T2010")

# If time is omitted, it will be set to midnight

parse_datetime("20101010")

parse_datetime("2010-10-01T2010")

# If time is omitted, it will be set to midnight

parse_datetime("20101010")


```

- **parse_date()** expects a four digit year, a - or /, the month, a - or /, then the day:

```{r, ex. 21}

parse_date("2010-10-01")

```

- **parse_time()** expects the hour, :, minutes, optionally : and seconds, and an optional am/pm specifier:

```{r, ex. 22}

library(hms)

parse_time("01:10 am")

parse_time("20:10:01")

```

Base R doesn’t have a great built in class for time data, so we use the one provided in the hms package.

If these defaults don’t work for your data you can supply your own date-time format, built up of the following pieces:




**Year**

- %Y (4 digits).
- %y (2 digits); 00-69 -> 2000-2069, 70-99 -> 1970-1999.

**Month**

- %m (2 digits).
- %b (abbreviated name, like “Jan”).
- %B (full name, “January”).

**Day**
- %d (2 digits).
- %e (optional leading space).

**Time**
- %H 0-23 hour.
- %I 0-12, must be used with %p.
- %p AM/PM indicator.
- %M minutes.
- %S integer seconds.
- %OS real seconds.
- %Z Time zone (as name, e.g. America/Chicago). Beware of abbreviations: if you’re American, note that “EST” is a Canadian time zone that does not have daylight savings time. It is not Eastern Standard Time! We’ll come back to this time zones.
- %z (as offset from UTC, e.g. +0800).
Non-digits
- %. skips one non-digit character.
- %* skips any number of non-digits.

The best way to figure out the correct format is to create a few examples in a character vector, and test with one of the parsing functions. For example:


If you’re using %b or %B with non-English month names, you’ll need to set the lang argument to locale(). See the list of built-in languages in date_names_langs(), or if your language is not already included, create your own with date_names().

```{r}
parse_date("22/02/21", "%m/%d/%y")

parse_date("22/02/21", "%d/%m/%y")

parse_date("22/02/21", "%y/%m/%d")

parse_date("2 janvier 2021", "%d %B %Y", locale = locale("fr"))
#
```


## 11.4 Parsing a file

Now that you’ve learned how to parse an individual vector, it’s time to return to the beginning and explore how readr parses a file. There are two new things that you’ll learn about in this section:

**1. How readr automatically guesses the type of each column.**

**2. How to override the default specification.**



### 11.4.1 Strategy

readr uses a heuristic to figure out the type of each column: it reads the first 1000 rows and uses some (moderately conservative) heuristics to figure out the type of each column. You can emulate this process with a character vector using **guess_parser()**, which returns readr’s best guess, and **parse_guess()** which uses that guess to parse the column:



```{r}
guess_parser("2010-10-01")

guess_parser("15:01")

guess_parser(c("TRUE", "FALSE"))

guess_parser(c("1", "5", "9"))

guess_parser(c("12,352,561"))

str(parse_guess("2010-10-10"))

```


The heuristic tries each of the following types, stopping when it finds a match:

- logical: contains only “F”, “T”, “FALSE”, or “TRUE”.

- integer: contains only numeric characters (and -).

- double: contains only valid doubles (including numbers like 4.5e-5).

- number: contains valid doubles with the grouping mark inside.

- time: matches the default time_format.

- date: matches the default date_format.

- date-time: any ISO8601 date.

If none of these rules apply, then the column will stay as a vector of strings.

### 11.4.2 Problems


These defaults don’t always work for larger files. There are two basic problems:

1. The first thousand rows might be a special case, and readr guesses a type that is not sufficiently general. For example, you might have a column of doubles that only contains integers in the first 1000 rows.

2. The column might contain a lot of missing values. If the first 1000 rows contain only NAs, readr will guess that it’s a logical vector, whereas you probably want to parse it as something more specific.

readr contains a challenging CSV that illustrates both of these problems:


```{r}
challenge <- read_csv(readr_example("challenge.csv"))


```

(Note the use of readr_example() which finds the path to one of the files included with the package)

There are two printed outputs: the column specification generated by looking at the first 1000 rows, and the first five parsing failures. It’s always a good idea to explicitly pull out the problems(), so you can explore them in more depth:

```{r}
problems(challenge)
```


```{r}
tail(challenge)
```


That suggests we need to use a date parser instead. To fix the call, start by copying and pasting the column specification into your original call:

```{r}
challenge <- read_csv(
  readr_example("challenge.csv"), 
  col_types = cols(
    x = col_double(),
    y = col_logical()))
```
Then you can fix the type of the y column by specifying that y is a date column:

```{r}
challenge <- read_csv(
  readr_example("challenge.csv"), 
  col_types = cols(
    x = col_double(),
    y = col_date()
  ))



tail(challenge)
```


I highly recommend always supplying col_types, building up from the print-out provided by readr. This ensures that you have a consistent and reproducible data import script. If you rely on the default guesses and your data changes, readr will continue to read it in. If you want to be really strict, use stop_for_problems(): that will throw an error and stop your script if there are any parsing problems.

### 11.4.3 Other strategies


There are a few other general strategies to help you parse files:

- In the previous example, we just got unlucky: if we look at just one more row than the default, we can correctly parse in one shot:

```{r}
challenge2 <- read_csv(readr_example("challenge.csv"), guess_max = 1001)

challenge2

```


- Sometimes it’s easier to diagnose problems if you just read in all the columns as character vectors:

```{r}
challenge2 <- read_csv(readr_example("challenge.csv"), 
  col_types = cols(.default = col_character()))
```

- This is particularly useful in conjunction with type_convert(), which applies the parsing heuristics to the character columns in a data frame.

```{r}
df <- tribble(
  ~x,  ~y,
  "1", "1.21",
  "2", "2.32",
  "3", "4.56"
)
df

type_convert(df)

```


If you’re reading a very large file, you might want to set n_max to a smallish number like 10,000 or 100,000. That will accelerate your iterations while you eliminate common problems.

If you’re having major parsing problems, sometimes it’s easier to just read into a character vector of lines with read_lines(), or even a character vector of length 1 with read_file(). Then you can use the string parsing skills you’ll learn later to parse more exotic formats.

### 11.5 Writing to a file


readr also comes with two useful functions for writing data back to disk: write_csv() and write_tsv(). Both functions increase the chances of the output file being read back in correctly by:

Always encoding strings in UTF-8.

Saving dates and date-times in ISO8601 format so they are easily parsed elsewhere.

If you want to export a csv file to Excel, use write_excel_csv() — this writes a special character (a “byte order mark”) at the start of the file which tells Excel that you’re using the UTF-8 encoding.

The most important arguments are x (the data frame to save), and path (the location to save it). You can also specify how missing values are written with na, and if you want to append to an existing file.

write_csv(challenge, "challenge.csv")
Note that the type information is lost when you save to csv:

```{r}
challenge

write_csv(challenge, "challenge-2.csv")
read_csv("challenge-2.csv")

```


This makes CSVs a little unreliable for caching interim results—you need to recreate the column specification every time you load in. There are two alternatives:


write_rds() and read_rds() are uniform wrappers around the base functions readRDS() and saveRDS(). These store data in R’s custom binary format called RDS:


```{r}
write_rds(challenge, "challenge.rds")
read_rds("challenge.rds")

```

The feather package implements a fast binary file format that can be shared across programming languages:


```{r}
library(feather)
write_feather(challenge, "challenge.feather")
read_feather("challenge.feather")
```

Feather tends to be faster than RDS and is usable outside of R. RDS supports list-columns (which you’ll learn about in many models); feather currently does not.

## 11.6 Other types of data


To get other types of data into R, we recommend starting with the tidyverse packages listed below. They’re certainly not perfect, but they are a good place to start. For rectangular data:

haven reads SPSS, Stata, and SAS files.

readxl reads excel files (both .xls and .xlsx).

DBI, along with a database specific backend (e.g. RMySQL, RSQLite, RPostgreSQL etc) allows you to run SQL queries against a database and return a data frame.

For hierarchical data: use jsonlite (by Jeroen Ooms) for json, and xml2 for XML. Jenny Bryan has some excellent worked examples at https://jennybc.github.io/purrr-tutorial/.

For other file types, try the R data import/export manual and the rio package.












    